---
title: "Coursera.EDA.Wk1.Assignment_1"
author: "MCCurcio"
date: "12/25/2020"
output: html_document
---

# Coursera-EDA-Wk1-Assignment 1

This assignment uses data from the [UC Irvine Machine Learning Repository](http://archive.ics.uci.edu/ml/index.php).

We are using the [Individual household electric power consumption Data Set](https://archive.ics.uci.edu/ml/datasets/individual+household+electric+power+consumption)

**The 9 variables in the dataset are taken from the UCI web site:**

- Date, Time, Global_active_power, Global_reactive_power, Voltage, Global_intensity, Sub_metering_1, Sub_metering_2, Sub_metering_3


**Tasks**

1. Use GitHub material from Roger Peng.  
    - Use: `git clone https://github.com/rdpeng/ExData_Plotting1`  
  
2. Save to MCCurcio GitHub repository?  
    
3. Produce 4 plots;  
    1. Histogram of Global_active_power  
    2. Plot of Global_active_power versus datetime  
    3. Plot  of Sub meterings versus datetime  
    4. 2 x 2 Plots  
    
---

## Data Preparation

### Load Household Power Consumption(hpc)

NOTE: The date column is read into `hpc` dataframe **with** date = 
      *col_date(format = "%d/%m/%Y")*. This makes subsetting data
      into Feb. 1 and 2 easier.
```{r message=FALSE, warning=FALSE}
library(readr)

setwd("~/Dropbox/R_exercises/Coursera/Exploratory Data Analysis/week1/ExData_Plotting1")

hpc <- read_delim("exdata_data_household_power_consumption.zip", 
                  ";", 
                  trim_ws = TRUE,
                  escape_double = FALSE, 
                  col_types = cols(Date = col_date(format = "%d/%m/%Y"),
                                   Time = col_character(),
                                   Global_intensity = col_skip()))

str(hpc)
```

### Subset Household Power Consumption(hpc) to Feb 1-2, 2007
Use between Dates: 2007-02-01 and 2007-02-02.
```{r}
btwn_two_dates <- function(x, y){
      hpc[hpc$Date >= x & hpc$Date <= y,]
      }

DATE1 <- as.Date("2007-02-01")
DATE2 <- as.Date("2007-02-02")

hpc_feb <- btwn_two_dates(DATE1, DATE2)

names(hpc_feb)
```

### Generate DateTime stamp for graphing.
```{r message=FALSE, warning=FALSE}
library(lubridate)

hpc_feb$Date <- as.character(hpc_feb$Date)
class(hpc_feb$Date)

hpc_feb$DateTime <- paste(hpc_feb$Date, hpc_feb$Time)
hpc_feb$DateTime <- parse_date_time(hpc_feb$DateTime, 'Ymd HMS')

head(hpc_feb$DateTime)
```

---

## Make 4 Plots 

### Plot 1 - `mcc_plot_1.png`
Histogram of `Global_active_power` vs frequency
```{r}
plot_1 <- function(){
    hist(hpc_feb$Global_active_power, 
         main = "Global Active Power",
         xlab = "Global Active Power (kW)",
         ylim = c(0, 1400), # Changed x & y limits to read values
         xlim = c(0, 8),
         col = "red")
         # Default for this graphic is width = 480, height = 480.
         # To change the size add the parameters width & height.
}

plot_1()

png(file = "mcc_plot_1.png")
plot_1()
dev.off()
```




### Plot 2 - `mcc_plot_2.png`
Plot `Global_active_power` versus `datetime`
```{r}
plot_2 <- function(){
    plot(x = hpc_feb$DateTime,
         y = hpc_feb$Global_active_power,
         type = "l",
         main = "Global Active Power vs Datetime",
         ylim = c(0, 8),
         ylab = "Global Active Power (kW)",
         xlab = "DateTime")
         # Default for this graphic is width = 480, height = 480.
         # To change the size add the parameters width & height.
}

plot_2()

png(file = "mcc_plot_2.png")
plot_2()
dev.off()
```


### Plot 3 - `mcc_plot_3.png`
Plot  of Sub meterings versus datetime
```{r}
plot_3 <- function(){
    with(hpc_feb, {
         plot(x = DateTime, 
         y = Sub_metering_1, 
         ylab = "Energy Sub Metering",
         type = "l",
         ylim = c(0, 40))
    lines(x = DateTime, 
          y = Sub_metering_2, 
          col = "red")
    lines(x = DateTime, 
          y = Sub_metering_3, 
          col = "blue")
    # add legend
	  legend("topright", 
	         c("Sub_metering_1","Sub_metering_2", "Sub_metering_3"),
	         col = c("black", "red", "blue"),
	         lty = c(1, 1, 1),
	         box.col = "white")
    }
    )
}

plot_3()

png(file = "mcc_plot_3.png")
plot_3()
dev.off()
```

### Plot 4 - `mcc_plot_4.png`
**plot_volts**
```{r}
plot_volts <- function(){
      plot(x = hpc_feb$DateTime,
           y = hpc_feb$Voltage,
           xlab = "DateTime",
           ylab = "Voltage",
           type = "l")
}

plot_volts()
```

**plot_grp**
```{r}
plot_grp <- function(){
      plot(x = hpc_feb$DateTime,
           y = hpc_feb$Global_reactive_power,
           xlab = "DateTime",
           ylab = "Global_reactive_power",
           type = "l")
}

plot_grp()
```


**2 x 2 Plots**
```{r}
plot_4 <- function(){
    par(mfrow = c(2, 2))
    plot_2()
    plot_volts()
    plot_3()
    plot_grp()
}

plot_4()

png(file = "mcc_plot_4.png")
plot_4()
dev.off()
```


